Analyzing Crime Trends in Toronto: Insights from the Police Annual Statistical Report

### 2024-09-27 paper/paper.qmd
Final Version

---
title: "Analyzing Crime Trends in Toronto: Insights from the Police Annual Statistical Report"
# subtitle:
author: 
  - Yizhe Chen
thanks: "Code and data are available at: https://github.com/YizheChenUT/STA304TermPaper1.git"
date: today
date-format: long
abstract: "This paper analyzes crime trends in Toronto using the Police Annual Statistical Report from 2014 to 2023. Data from Open Data Toronto is used to evaluate crime distributions across different categories and regions. Analysis reveals that some areas, particularly in divisions such as D13 and D32, report the highest number of property-related crimes like auto theft. Suburban areas also illustrate rising trends in specific crime subtypes. These findings can help optimize policing strategies and public safety resource allocations."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

@@

# Data {#sec-data}
The data used in this analysis was obtained from the Open Data Toronto portal [@OpenDataDataset]. The data set includes reported incidents categorized by type, division, and year across Toronto, covering the period from 2014 to 2023. The data collection is a key resource for tracking public safety, and its availability promotes transparency in policing efforts. However, the dataset also raises ethical concerns, particularly around data privacy. Although anonymized, the geospatial breakdown could still enable the identification of patterns in specific communities, potentially stigmatizing certain areas.

@@

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-Crimes_by_Div
#| fig-cap: Crimes by Division and Crime Type

## Create a bar chart comparing crime types and incidents by division
## Citation: https://allisonhorst.github.io/palmerpenguins/
## Citation: https://ggplot2.tidyverse.org
crime_by_division_plot %>% 
  ggplot(aes(y = reorder(division_name, -total_incidents), x = total_incidents, fill = crime_subtype)) +
  geom_bar(stat = "identity", position = "dodge") +  # Using a bar chart for comparison
  labs(
    title = "Crime Incidents by Division and Crime Type in Toronto",
    y = "Division",
    x = "Total Incidents",
    fill = "Crime Type"
  ) +
  theme_minimal()  # Minimal theme for clean visuals

@@

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-Clearance_Rate
#| fig-cap: Clearance Rate by Crime Type

## Create a grouped bar chart to show clearance rate by crime type and division
## Citation: https://allisonhorst.github.io/palmerpenguins/
## Citation: https://ggplot2.tidyverse.org
clearance_rate_plot %>%
  ggplot(aes(y = reorder(division_name, -avg_clearance_rate), x = avg_clearance_rate, fill = crime_subtype)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +  # Grouped bar chart for clearance rates
  labs(
    title = "Clearance Rate by Crime Type and Division in Toronto",
    y = "Division",
    x = "Average Clearance Rate",
    fill = "Crime Type"
  ) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +  # Convert y-axis to percentage
  theme_minimal()  # Minimal theme for clean visuals

>>>>

---
title: "Analyzing Crime Trends in Toronto: Insights from the Police Annual Statistical Report"
# subtitle:
author: 
  - Yizhe Chen
thanks: "Code and data are available at: https://github.com/YizheChenUT/Analyzing-Crime-Trends-in-Toronto-Insights-from-the-Police-Annual-Statistical-Report.git"
date: today
date-format: long
abstract: "This paper analyzes crime trends in Toronto using the Police Annual Statistical Report from 2014 to 2023. Data from Open Data Toronto is used to evaluate crime distributions across different categories and regions. Analysis reveals that some areas, particularly in divisions such as D13 and D32, report the highest number of property-related crimes like auto theft. Suburban areas also illustrate rising trends in specific crime subtypes. These findings can help optimize policing strategies and public safety resource allocations."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

@@

# Data {#sec-data}
The data used in this analysis was obtained from the Open Data Toronto portal [@OpenDataDataset]. Data was cleaned and analyzed using the language R [@citeR]. The data set includes reported incidents categorized by type, division, and year across Toronto, covering the period from 2014 to 2023. The data collection is a key resource for tracking public safety, and its availability promotes transparency in policing efforts. However, the dataset also raises ethical concerns, particularly around data privacy. Although anonymized, the geospatial breakdown could still enable the identification of patterns in specific communities, potentially stigmatizing certain areas.

@@

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-Crimes_by_Div
#| fig-cap: Crimes by Division and Crime Type
#| fig-width: 10
#| fig-height: 8

## Create a bar chart comparing crime types and incidents by division
## Citation: https://allisonhorst.github.io/palmerpenguins/
## Citation: https://ggplot2.tidyverse.org
crime_by_division_plot %>% 
  ggplot(aes(y = reorder(division_name, -total_incidents), x = total_incidents, fill = crime_subtype)) +
  geom_bar(stat = "identity", position = "dodge") +  # Using a bar chart for comparison
  labs(
    title = "Crime Incidents by Division and Crime Type in Toronto",
    y = "Division",
    x = "Total Incidents",
    fill = "Crime Type"
  ) +
  theme_minimal() + # Minimal theme for clean visuals
  theme(
    axis.text.y = element_text(size = 8, margin = margin(r = 10)), # Add the spacing of y-axis
    plot.title = element_text(hjust = 0.5), # Placing title on the middle
    legend.position = "bottom"  # Placing legend on the bottom
  ) +
  guides(fill = guide_legend(ncol = 2))  # Making the legend 2 columns

@@

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-Clearance_Rate
#| fig-cap: Clearance Rate by Crime Type
#| fig-width: 10
#| fig-height: 8

## Create a grouped bar chart to show clearance rate by crime type and division
## Citation: https://allisonhorst.github.io/palmerpenguins/
## Citation: https://ggplot2.tidyverse.org
clearance_rate_plot %>%
  ggplot(aes(y = reorder(division_name, -avg_clearance_rate), x = avg_clearance_rate, fill = crime_subtype)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +  # Grouped bar chart for clearance rates
  labs(
    title = "Clearance Rate by Crime Type and Division in Toronto",
    y = "Division",
    x = "Average Clearance Rate",
    fill = "Crime Type"
  ) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +  # Convert y-axis to percentage
  theme_minimal() +  # Minimal theme for clean visuals
  theme(
    axis.text.y = element_text(size = 12, margin = margin(r = 10)), # Add the spacing of y-axis
    plot.title = element_text(hjust = 0.5), # Placing title on the middle
    legend.position = "bottom"  # Placing legend on the bottom
  ) +
  guides(fill = guide_legend(ncol = 2))  # Making the legend 2 columns
